The imprint of global magma oceans on exoplanet demographics

### 2023-03-10 src/tex/ms.tex
clean up density measurements

\subsection{Measurement uncertainties}
Under real-world conditions, the planetary properties in question can only be probed with a finite precision that is specific to each exoplanet mission.
%\subsubsection{PLATO}
\plato\ (PLAnetary Transits and Oscillation of stars) is an ESA mission designed to characterize terrestrial planets in the habitable zones of Sun-like stars, a goal to be achieved via long-term high-precision photometric monitoring of a sample of bright stars~\citep{Rauer2016}.
\plato's definition study report~\citep{plato2017} states precision requirements for planet radii~(\SI{3}{\percent}), planet masses through radial velocity~(RV) follow-up~(\SI{10}{\percent}), and stellar masses, radii, and ages~(\SI{10}{\percent}).
For uncertainties in bulk density measurements, we propagate the errors of the mass measurements assuming $\sigma_\mathrm{M_\mathrm{P}} = \SI{10}{\percent}$.
We base our assessment of \plato's diagnostic power on these estimates.

\subsection{Detection bias, sample selection, and sensitivity}\label{sec:sensitivity} % Turbet2019 lists precision for CHEOPS, PLATO?, TESS?
Not all transiting planets are detectable with the same likelihood and detection biases have an impact on the demographic measurements we are interested in.
A detailed characterization of the detection biases of individual missions would not be justifiable given the uncertainties of the theoretical predictions.
Instead, we derived generic observing limits that reflect the limitations of state-of-the-art transit surveys.

@@

\subsubsection{Additional planet mass measurements}\label{sec:res_followup}
The dependencies of a detection change when additional information beyond planet radii is available for the characterized planet population.
Since the planet bulk density $\rho \propto R^{-3}$, the signal is stronger when measured in bulk density instead of radius:
With an optimistic choice of geophysical parameters (see above), the average measured radius change is \SI{+15}{\percent} whereas the average density change is \SI{-33}{\percent}.
Constraining bulk densities requires planetary mass measurements.
We now examine the value of a follow-up campaign to constrain planet masses with respect to the detectability of the runaway greenhouse threshold.

As can be seen in Fig.~\ref{fig:plato_fgrid}, the detectability is enhanced when radius constraints can be augmented with mass measurements: At unchanged sample size, the difference in Bayesian evidence can be up to an order of magnitude larger.
Consequently, we achieve a statistically significant detection with smaller samples or lower dilution factors $f_\mathrm{rgh}$.
A density-based survey of 100 targets is roughly equivalent to a radius-based survey of \var{N_plato} targets.
At $N=100$, pure radius measurements require $f_\mathrm{rgh} \gtrsim 0.4$ whereas bulk density measurements enable a detection from $f_\mathrm{rgh} \gtrsim 0.1$.

@@

\subsubsection{The value of follow-up campaigns}
As runaway greenhouse phases leave a stronger imprint in bulk density than in planet radius (see Sect.~\ref{sec:res_followup}), it would be beneficial to obtain constraints on planetary masses and test the runaway greenhouse hypothesis in density space instead of radius space.
This way, useful constraints can be obtained under more pessimistic conditions, e.g., a low predominant water content of planetary surfaces and atmospheres or a smaller available planet sample.
For a mission design similar to PLATO, a density-based hypothesis test on about a third of the overall sample is equivalent to a pure radius-based analysis.
At a fixed sample size, key parameters of the runaway greenhouse models can be more narrowly constrained when additional mass measurements are available.

>>>>

\subsection{Measurements and their uncertainties}
Under real-world conditions, the planetary properties in question can only be probed with a finite precision that is specific to each exoplanet mission.
%\subsubsection{PLATO}
\plato\ (PLAnetary Transits and Oscillation of stars) is an ESA mission designed to characterize terrestrial planets in the habitable zones of Sun-like stars, a goal to be achieved via long-term high-precision photometric monitoring of a sample of bright stars~\citep{Rauer2016}.
\plato's definition study report~\citep{plato2017} states precision requirements for planet radii~(\SI{3}{\percent}), planet masses through radial velocity~(RV) follow-up~(\SI{10}{\percent}), and stellar masses, radii, and ages~(\SI{10}{\percent}).
We adopted these estimates and assumed a \SI{10}{\percent} error on instellation measurements.

Since planetary bulk density $\rho \propto R_\mathrm{P}^{-3}$, we expect a stronger runaway greenhouse signal when measured through bulk density instead of radius.
We thus simulated measurements of planetary densities assuming the mass-radius relation defined above.
For uncertainties in bulk density measurements, we propagated the errors of the mass measurements assuming $\sigma_\mathrm{M_\mathrm{P}} = \SI{10}{\percent}$.

\subsection{Detection bias, target selection, and sensitivity}\label{sec:sensitivity} % Turbet2019 lists precision for CHEOPS, PLATO?, TESS?
Not all transiting planets are detectable with the same likelihood and detection biases have an impact on the demographic measurements we are interested in.
A detailed characterization of the detection biases of individual missions would not be justifiable given the uncertainties of the theoretical predictions.
Instead, we derived generic observing limits that reflect the limitations of state-of-the-art transit surveys.

@@

\subsubsection{Additional planet mass measurements}\label{sec:res_followup}
Comparing a measurement of the signal in radius space with a measurement in density space (which requires planetary mass measurements), a stronger detection occurs in the latter case:
With an optimistic choice of geophysical parameters (see above), the average measured radius change is \SI{+15}{\percent} whereas the average density change is \SI{-33}{\percent}.

%We now examine the value of a follow-up campaign to constrain planet masses with respect to the detectability of the runaway greenhouse threshold.
As can be seen in Fig.~\ref{fig:plato_fgrid}, the detectability is enhanced when radius constraints can be augmented with mass measurements: At unchanged sample size, the difference in Bayesian evidence can be up to an order of magnitude larger.
Consequently, we achieve a statistically significant detection with smaller samples or lower dilution factors $f_\mathrm{rgh}$.
A density-based survey of 100 targets is roughly equivalent to a radius-based survey of \var{N_plato} targets.
At $N=100$, pure radius measurements require $f_\mathrm{rgh} \gtrsim 0.4$ whereas bulk density measurements enable a detection from $f_\mathrm{rgh} \gtrsim 0.1$.

@@

\subsubsection{The value of follow-up campaigns}
The dependencies of a detection change when additional information beyond planet radii is available for the characterized planet population.
As runaway greenhouse phases leave a stronger imprint in bulk density than in planet radius (see Sect.~\ref{sec:res_followup}), it would be beneficial to obtain constraints on planetary masses and test the runaway greenhouse hypothesis in density space instead of radius space.
This way, useful constraints can be obtained under more pessimistic conditions, e.g., a low predominant water content of planetary surfaces and atmospheres or a smaller available planet sample.
For a mission design similar to PLATO, a density-based hypothesis test on about a third of the overall sample is equivalent to a pure radius-based analysis.
At a fixed sample size, key parameters of the runaway greenhouse models can be more narrowly constrained when additional mass measurements are available.
